Long-form recording of infant body position in the home using wearable inertial sensors

### 2023-09-08 rev1-manuscript/manuscript.tex
revision

\abstract{%
Long-form audio recordings have had a transformational effect on the study of infant language acquisition by using mobile, unobtrusive devices to gather full-day, real-time data that can be automatically scored. How can we produce similar data in service of measuring infants' everyday motor behaviors, such as body position? The aim of the current study was to validate long-form recordings of infant position (supine, prone, sitting, upright, held by caregiver) based on machine learning classification of data from inertial sensors worn on infants' ankles and thighs. Using over 100 hours of video recordings synchronized with inertial sensor data from infants in their homes, we demonstrate that body position classifications are sufficiently accurate to measure infant behavior. Moreover, classification remained accurate when predicting behavior later in the session when infants and caregivers were unsupervised and went about their normal activities, showing that the method can handle the challenge of measuring unconstrained, natural activity. Next, we show that the inertial sensing method has convergent validity by replicating age differences in body position found using other methods with full-day data captured from inertial sensors. We end the paper with a discussion of the novel opportunities that long-form motor recordings afford for understanding infant learning and development.
}

@@

In this paper, we present an inertial sensing method to classify infants' full-day, real-time body position. Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences. We begin by identifying the key features of wearable audio recorders that should be replicated in long-form recordings of motor behavior. Next, we review the current state-of-the-art in measuring infant motor behavior---video and survey data---and their limitations in capturing real-time, full-day behavior. Finally, we discuss the advantages of using inertial sensing to classify motor behavior. Despite promising past results in brief, supervised sessions \autocite{FranchakScott2021,AiraksinenRasanen2020,GreenspanCunha2021,AiraksinenGallen2022}, the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings in what we term a \emph{distal comparison}.

@@

However, automatic classification can only replace human annotation if it is sufficiently accurate. An independent assessment of the LENA® algorithms found mixed results about classification accuracy. Correlations between human transcribed counts of adult words and child vocalizations against LENA®'s automatic counts were strong, \emph{r} = .698 and \emph{r} = .649, respectively \autocite{CristiaLavechin2020}. For other metrics, such as the number of ``conversational turns'' between the child and communicative partners, agreement was poor (\emph{r} = .364). Thus, for some use cases (and for some metrics), long-form audio recordings provide a mobile, unobtrusive way to automatically score real-time data over a full day.

@@

Video and survey methods are the current state-of-the-art in assessing infants' gross motor behavior in naturalistic tasks. Although each method has complementary advantages and disadvantages for characterizing infants' everyday motor experiences, neither method fulfills all five key features of long-form audio recordings reviewed in the previous section.

Video observation is the most common way of measuring infant motor behavior in home recordings. Most often, an experimenter with a handheld camera follows infants from room to room to ensure that their movements are visible throughout the recording session \autocite{Karasik2011,ChenSchneider2022,HerzbergFletcher2021}. The primary advantage of video recording is that it captures real-time behavior. Standard 30 Hz video recording is adequate to capture changes in infant body position that occur on the timescale of seconds. However, requiring an experimenter to operate a camera is obtrusive, whereas relying on a stationary camera means that infants will be absent from view as they move from place to place. Moreover, video observation cannot easily scale to long durations or large numbers of participants. An experimenter cannot follow behind infants to record their behavior from morning to night; typical video recording sessions last 45-120 minutes \autocite{Karasik2011,ChenSchneider2022,HerzbergFletcher2021}, far short of capturing the variety of activities across the full daily routine. Even if full-day videos were available, the lack of suitable automatic classification tools means that the human cost of annotation would be immense. Our annotation of body position takes approximately 2-5 hours to complete for every hour of video (depending on how often infants switch positions), meaning that a full ``waking day'' of approximately 11 hours for a 12-month-old \autocite{GallandTaylor2012} could take 22-55 hours of labor to annotate.

In contrast, survey methods such as daily diaries/inventories or ecological momentary assessment (EMA) are mobile, unobtrusive, can be applied across an entire day, and do not need laborious annotation. Diary studies provide caregivers with logs or structured interviews to record activities \autocite{MajnemerBarr2005,KarasikKuchirko2022}. Ecological momentary assessment uses smartphone notifications to prompt caregivers to make repeated estimates of behaviors throughout the day \autocite{Survey,KadookaCaufield2021}. Although such survey responses are valuable in aggregate, they lack the real-time temporal resolution to describe moment-to-moment changes in behavior. At best, EMA surveys prompt caregivers to make hourly observations; increasing the number of surveys per day would be too burdensome for the respondent. Thus, despite being a useful tool for estimating broad developmental changes and individual differences in infants' motor experiences, survey methods are not suited for capturing within-participant temporal dynamics.

@@

Measuring infant movement with inertial movement units (IMUs) is a promising avenue for long-form recordings of motor behavior in the home \autocite{CliffReilly2009,Barbaro2019,BruijnsTruelove2020,LoboHall2019}. Lightweight sensors (10-30 g) can be embedded in garments to make recordings fully \emph{mobile}, and they are \emph{unobtrusive} because they do not require a researcher to follow with a camcorder. Many commercially-available IMUs have \textgreater{} 12 hour battery life with onboard storage to record \emph{real-time}, \emph{full-day} motion data at a high sampling rate (e.g., 50-100 Hz).

The open question is whether \emph{automatic classification} is sufficiently accurate to measure movement categories that are relevant to developmental and clinical research, and whether measurements continue to be accurate over long recording periods. Data processing algorithms are needed to classify the raw sensor data (i.e., linear and angular acceleration time series) into meaningful categories. Categorizing body position---supine on the back, prone on the belly, sitting, upright, or held off the ground by a caregiver---is complex because movement can vary greatly \emph{within} a body position. An upright infant can be standing still or can be walking briskly across the room. A prone infant can be stationary in ``tummy time'', or they can crawl in a myriad of ways \autocite{AdolphVereijken1998}. Moreover, the configuration of the arms, legs, and torso within a body position can vary greatly in everyday contexts. Infants can sit on the floor in a tripod position with support from an arm, in a ``V'' position with legs fully extended, or in a ``W'' position with knees bent. Sitting on a caregiver's lap without the need to maintain balance means that the legs can dangle and the torso can lean in different directions. Sitting in a high chair or car seat likely restricts the range of torso orientations compared with sitting independently or on a caregiver's lap. Finally, caregivers frequently pick up and transport infants, creating motion signals that need to be differentiated from independent activity \autocite{KwonZavos2019,PatelShi2019}.

Modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults \autocite{Preece2009,Arif2015}, children \autocite{NamPark2013,RenDing2016,StewartNarayanan2018}, and infants \autocite{YaoPlotz2019,FranchakScott2021,AiraksinenRasanen2020}. Three prior investigations have used different machine learning techniques to categorize infant body position from multiple IMUs towards the goal of collecting full-day data. \textcite{AiraksinenRasanen2020} tested 4- to 8-month-olds in a laboratory visit with a 4-sensor array (one on each thigh and one below each shoulder), and found 95\% accuracy in distinguishing between body position categories that crawling infants could perform on the floor (excluding times that infants were held by caregivers). Using a wider age range of 6-18 months, \textcite{FranchakScott2021} found 98\% accuracy (\emph{kappa} = 95\%) in a laboratory validation study with a 3-sensor array (ankle, knee, thigh on a single leg) in categorizing body position that included infants who could both crawl and walk and also included a category for caregiver holding. Most recently, \textcite{AiraksinenGallen2022} conducted a validation study of body position classification in either a home or clinic testing 4- to 19-month-olds with a 4-sensor system, refining their previous method to detect moments that infants were carried by caregivers. Classification accuracy did not vary between lab and home settings, and was generally high (95\%, \emph{kappa} = .93). In contrast to the three machine learning studies that used 3-4 sensors \autocite{AiraksinenRasanen2020,FranchakScott2021,AiraksinenGallen2022}, \textcite{GreenspanCunha2021} used orientation from a single hip-worn sensor to measure body position with a high degree of accuracy (\emph{kappa} = .84). Although all four studies yielded promising classification accuracy, accuracy was assessed in brief (15-60 minute) sessions supervised by a researcher, leaving the open question of how well body position classification will scale to testing across an entire day of natural home life.

@@

Accordingly, the overarching goal of the current study is to test the validity of long-form body position recording in the home during unsupervised, everyday behavior. Supervised recordings from past work \autocite{FranchakScott2021,AiraksinenRasanen2020,AiraksinenGallen2022}, whether in the home or in the lab, let researchers set up the situation to encourage or restrict certain behaviors. Prior work focused on ``free play'', in which caregivers were asked to play with the infant without restraining the infant or otherwise shaping how they could move. However, in a real day, non-play activities (e.g., eating lunch in a high chair) create challenging situations for applying automated classification of body position. For example different types of sitting---independently on the floor, supported on a caregiver's lap, or restrained in a high chair---all need to be scored as sitting. Moreover, classifiers must be able to detect new variants of behaviors that might arise over the course of a real day; it is impossible for researchers to gather training data for every possible variation that might occur. Thus, our central question is whether models trained on video-recorded observations at the beginning of the day generalize to predict behavior at a later time. Assessing the validity of temporally \emph{distal} periods---when infants and caregivers are unsupervised and free to follow their everyday routines---is a crucial step to establish whether automatic classification can be used to measure body position across a day.

@@

Infants were recruited in one of two age groups: \emph{Younger} infants were between 4 and 7 months and \emph{older} infants were between 11 and 14 months. There were 8 infants in the 4-7 month group and 14 in the 11-14 month group. Ten infants were female and 12 were male. Families were recruited through social media advertisements and from community events in Southern California. Most infants were reported to be either Hispanic and White (\emph{n} = 8) or Non-Hispanic and White (\emph{n} = 7). Families were compensated \$30 for every home recording session they completed. The University of California, Riverside Institutional Review Board reviewed and approved all procedures associated with the study. All caregivers gave their informed consent before the start of the study.

Most participants were tested in a single session (\emph{n} = 15), but 7 participants contributed between 2-4 sessions as part of an ongoing longitudinal study (3 from the younger group and 4 from the older group). Although including more data from some participants could over-represent their characteristics in the model, we reasoned that this drawback was outweighed by having more available data to use for training. Only 1 session was excluded due to a technical error---one of the four IMU sensors failed to record, resulting in an unusable set of data for classification. Across the two age groups, we report data on a total of 34 sessions, with 14 sessions from younger infants and 20 sessions from older infants. Across sessions, younger infants' age ranged from 3.8 to 7.2 months (\emph{M} = 5.2) and older infants' age ranged from 10.7 to 14.2 months (\emph{M} = 11.7). The total number of recording sessions (34) exceeded the number of sessions employed in comparable past work: 10 in \textcite{NamPark2013}, 15 in \textcite{FranchakScott2021}, 22 in \textcite{AiraksinenRasanen2020}, 23 in \textcite{GreenspanCunha2021}, and 33 in \textcite{YaoPlotz2019}.

@@

Four inertial movement units (IMUs) were used to record infant movement across the day (MC10 Biostamp). A custom garment was designed to hold the IMUs (Figure \ref{fig:garment}). Internal pockets were sewn into snug-fitting infant leggings so that IMUs would stay close to the body (reducing vibration). A pocket over the thigh and a pocket just above the ankle were sewn on the lateral surface of the right and left legs of the garment. Multiple sizes of the garment were created (modeled on US 3-6 mo, 6-9 mo, 9-12 mo, and 12-18 mo sizing). Caregivers indicated in advance which size would fit their infant, and to ``size down'' if between sizes to ensure a snug fit and minimize sensor movement. Each garment had a distinct pattern on the seat of the pants so that caregivers could identify front versus back and place the garment in the correct orientation.

Each sensor had sufficient battery and onboard storage to record accelerometer and gyroscope data for approximately 12 hours. We chose a sampling rate of 62.5 Hz (one of the available presets) based on prior work that used rates of 50-64 Hz \autocite{FranchakScott2021,AiraksinenRasanen2020,YaoPlotz2019}. Infants also wore a LENA® recorder throughout the day in the front pocket of a LENA® shirt, located near the infant's chest, to determine whether data could be simultaneously recorded from the LENA® and IMU sensors (LENA® data were not analyzed in the current study).

@@

Because the camera only had the battery life to record for \textasciitilde3 hours (automatically split into two 90-minute video files), this divided the day into different periods for analysis. As seen in the bottom of Figure \ref{fig:exemplar-timeline}, the \emph{video period} comprised the first three hours of recording starting from the researcher's arrival when they turned on the camera. The first 90-minute video file, termed the \emph{proximal comparison}, contained the activities during the guided call followed by a period during which infants and caregivers resumed their normal activities. Because this video contained the synchronization point, the data in this period had high temporal synchrony between IMU and video data. Synchronization errors were estimated to be less than 30-60 ms (1-2 video frames). The second video file comprised the \emph{distal comparison}. This video recorded the next 90 minutes of natural activity. However, because of a limitation in the camera, there was a variable gap of \textasciitilde40 s between the two videos, so synchronization in the distal comparison video was coarser, with estimated temporal offsets of \textasciitilde5 s in either direction.

@@

The proximal and distal comparison videos for each participant were annotated by trained human coders to classify infant \emph{body position} into one of 5 mutually-exclusive categories following the definitions in prior work \autocite{FranchakScott2021}: supine, prone, sitting, upright, or held by caregiver. All coding was done using Datavyu software (datavyu.org). A Databrary repository contains the entire video recording, coding files, and raw IMU data for a single participant (\url{https://nyu.databrary.org/volume/1580}).

Supine was coded when the infant was lying on their back, on their side, or was reclined up to a 45 degree angle. Prone was coded when the infant was lying on their stomach, was stationary supported by the hands/knees or the hands/feet, or was crawling. We scored sitting to include any form of the following seated positions: 1) infants sat with their buttocks on a surface, such as on the floor or a caregiver's lap, 2) infant was in a kneeling-sit position, in which their knees were on the ground with their legs tucked underneath the buttocks, and 3) infant was in a seating device, such as a high chair, that kept the torso oriented perpendicular to the ground (a reclined position, such as in a young infant's car seat, would be counted as supine). Upright was coded when the infant was standing or squatting on the ground with two feet or walking (regardless of whether infants' balance was assisted by a caregiver or with their hands holding onto something for support). Our goal in creating a category for ``held by caregiver'' was to separate times when infants were in control of their body position from times when they were suspended in the air (rather than resting on furniture or a surface). Held was coded when infants were carried off of the ground. However, when the caregiver was sitting with the infant in their lap the infant's body position was coded as if the caregiver was a surface (e.g., if the infant was sitting on the caregiver's lap this was coded as sitting). Times during the video when the infant was out of view were excluded. Periods when the sensor garment was adjusted or taken off the infant were also excluded, as were transitions between body positions.

@@

Classification training and prediction was conducted on a windowed dataset that summarized the raw, 62.5 Hz motion signals within 4-s windows. Data were reduced in time by creating overlapping moving windows (4-s long, comprising 250 samples) starting each second, which is a common unit of analysis in prior studies of human activity classification \autocite{FranchakScott2021,NamPark2013,AiraksinenRasanen2020}. For each 4-s window, we aggregated the 250 samples to create single scores for a variety of motion features---summary statistics that could be fed into the machine learning model. The minimum, maximum, 25th percentile, 75th percentile, mean, median, skew, kurtosis, standard deviation, and sum were computed for each signal (e.g., right thigh linear acceleration along the X-axis, left ankle pitch angular acceleration). The 10 summary statistics and 24 sensor signals generated 240 columns of motion features that described movement within each window. Furthermore, a series of cross-sensor and cross-orientation summaries (such as the correlation, magnitude, and difference between pairs of sensors) added an additional 196 columns of motion features. The 436 total motion features corresponded to a single body annotation code for each 4-s window. Windows were only used for training/testing if they contained a single body position for \textgreater{} 75\% (3 s) of time within the window to ensure that motion signals could be linked to an unambiguous example of each behavior.

@@

\caption{Metrics of agreement between human-annotated body position and model predictions of body position from the proximal comparison period. Overall accuracy (A) and Cohen's Kappa (B) are plotted separately for group models and individual models. Each blue circle represents the accuracy for each recording session. Horizontal black bars indicate the mean across sessions.}\label{fig:metrics}
\end{figure}

@@

\begin{tabular}{lllllll}
\toprule
 & \multicolumn{3}{c}{Group} & \multicolumn{3}{c}{Individual} \\
\cmidrule(r){2-4} \cmidrule(r){5-7}
Sensors & Median & Mean & SD & Median & Mean & SD\\
\midrule
All & 0.894 & 0.846 & 0.131 & 0.933 & 0.916 & 0.072\\ \midrule
Left Thigh/Ankle & 0.891 & 0.824 & 0.177 & 0.915 & 0.901 & 0.071\\
Right Thigh/Ankle & 0.842 & 0.800 & 0.167 & 0.897 & 0.898 & 0.086\\
Both Thighs & 0.859 & 0.791 & 0.195 & 0.907 & 0.898 & 0.072\\
Both Ankles & 0.797 & 0.719 & 0.196 & 0.864 & 0.844 & 0.108\\ \midrule
Left Thigh & 0.842 & 0.762 & 0.208 & 0.889 & 0.872 & 0.077\\
Left Ankle & 0.717 & 0.667 & 0.208 & 0.851 & 0.825 & 0.117\\
Right Thigh & 0.846 & 0.772 & 0.178 & 0.896 & 0.876 & 0.092\\
Right Ankle & 0.704 & 0.683 & 0.160 & 0.830 & 0.819 & 0.114\\
\bottomrule
\end{tabular}

@@

The first set of results showed that group and individual models trained from data during the proximal comparison period were accurate during the proximal period. This suggests that the immediate accuracy of body position predictions early in the recording session was strong, but does not address how accurately predictions will be later in the recording. In the next analysis, we examine long-term performance by testing how accurately models trained from the proximal period could predict body position during the distal comparison period. A single group model was created using all sessions' proximal period training data (rather than group models leaving out a single session); the same individual models were used. Distal videos had only coarse temporal synchrony with motion recordings which precluded calculating accuracy based on the proportion of matching events. Instead, we summed the amount of time infants were predicted to be in each of the 5 body position categories from the model and compared that to the summed time for the body positions based on human coding \autocite{YaoPlotz2019,FranchakScott2021}.

@@

\caption{Age differences in daily body position predicted from (A) group models and (B) individual models. Each circle represents one full-day recording session's proportion of time in each body position (y-axis) against age in months (x-axis). Black lines indicate best fit regression lines, which show decreases in supine time and increases in sitting and upright time with age.}\label{fig:age}
\end{figure}

>>>>

\abstract{%
Long form audio recordings have had a transformational effect on the study of infant language acquisition by using mobile, unobtrusive devices to gather full-day, real-time data that can be automatically scored. How can we produce similar data in service of measuring infants' everyday motor behaviors, such as body position? The aim of the current study was to validate long form recordings of infant position (supine, prone, sitting, upright, held by caregiver) based on machine learning classification of data from inertial sensors worn on infants' ankles and thighs. Using over 100 hours of video recordings synchronized with inertial sensor data from infants in their homes, we demonstrate that body position classifications are sufficiently accurate to measure infant behavior. Moreover, classification remained accurate when predicting behavior later in the session when infants and caregivers were unsupervised and went about their normal activities, showing that the method can handle the challenge of measuring unconstrained, natural activity. Next, we show that the inertial sensing method has convergent validity by replicating age differences in body position found using other methods with full-day data captured from inertial sensors. We end the paper with a discussion of the novel opportunities that long form motor recordings afford for understanding infant learning and development.
}

@@

In this paper, we present an inertial sensing method to classify infants' full-day, real-time body position. Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences. We begin by identifying the key features of wearable audio recorders that should be replicated in long-form recordings of motor behavior. Next, we review the current state-of-the-art in measuring infant motor behavior---video and survey data---and their limitations in capturing real-time, full-day behavior. Finally, we discuss the advantages of using inertial sensing to classify motor behavior. Despite promising past results in brief, supervised sessions \autocite{FranchakScott2021,AiraksinenRasanen2020,AiraksinenGallen2022}, the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings in what we term a \emph{distal comparison}.

@@

But automatic classification can only replace human annotation if it is sufficiently accurate. An independent assessment of the LENA® algorithms found mixed results about classification accuracy. Correlations between human transcribed counts of adult words and child vocalizations against LENA®'s automatic counts were strong, \emph{r} = .698 and \emph{r} = .649, respectively \autocite{CristiaLavechin2020}. For other metrics, such as the number of ``conversational turns'' between the child and communicative partners, agreement was poor (\emph{r} = .364). Thus, for some use cases (and for some metrics), long form audio recordings provide a mobile, unobtrusive way to automatically score real-time data over a full day.

@@

Video and survey methods are the current state-of-the-art in assessing infants' gross motor behavior in naturalistic tasks. Although each method has complementary advantages and disadvantages for characterizing infants' everyday motor experiences, neither method fulfills all five key features of long form audio recordings reviewed in the previous section.

Video observation is the most common way of measuring infant motor behavior in home recordings. Most often, an experimenter with a handheld camera follows infants from room to room to ensure that their movements are visible throughout the recording session \autocite{Karasik2011,ChenSchneider2022,HerzbergFletcher2021}. The primary advantage of video recording is that it captures real-time behavior.Standard 30 Hz video recording is adequate to capture changes in infant body position that occur on the timescale of seconds. However, requiring an experimenter to operate a camera is obtrusive, whereas relying on a stationary camera means that infants will be absent from view as they move from place to place. Moreover, video observation cannot easily scale to long durations or large numbers of participants. An experimenter cannot follow behind infants to record their behavior from morning to night; typical video recording sessions last 45-120 minutes \autocite{Karasik2011,ChenSchneider2022,HerzbergFletcher2021}, far short of capturing the variety of activities across the full daily routine. Even if full-day videos were available, the lack of suitable automatic classification tools means that the human cost of annotation would be immense. Our annotation of body position takes approximately 2-5 hours to complete for every hour of video (depending on how often infants switch positions), meaning that a full ``waking day'' of approximately 11 hours for a 12-month-old \autocite{GallandTaylor2012} could take 22-55 hours of labor to annotate.

In contrast, survey methods such as daily diaries/inventories or ecological momentary assessment (EMA) are mobile, unobtrusive, can be applied across an entire day, and do not need laborious annotation. Diary studies provide caregivers with logs or structured interviews record activities \autocite{MajnemerBarr2005,KarasikKuchirko2022}. Ecological momentary assessment uses smartphone notifications to prompt caregivers to make repeated estimates of behaviors throughout the day \autocite{Survey,KadookaCaufield2021}. Although such survey responses are valuable in aggregate, they lack the real-time temporal resolution to describe moment-to-moment changes in behavior. At best, EMA surveys prompt caregivers to make hourly observations; increasing the number of surveys per day would be too burdensome for the respondent. Thus, despite being a useful tool for estimating broad developmental changes and individual differences in infants' motor experiences, survey methods are not suited for capturing within-participant temporal dynamics.

@@

Measuring infant movement with inertial movement units (IMUs) is a promising avenue for long form recordings of motor behavior in the home \autocite{CliffReilly2009,Barbaro2019,BruijnsTruelove2020,LoboHall2019}. Lightweight sensors (10-30 g) can be embedded in garments to make recordings fully \emph{mobile}, and they are \emph{unobtrusive} because they do not require a researcher to follow with a camcorder. Many commercially-available IMUs have \textgreater{} 12 hour battery life with onboard storage to record \emph{real-time}, \emph{full-day} motion data at a high sampling rate (e.g., 50-100 Hz).

The open question is whether \emph{automatic classification} is sufficiently accurate to measure movement categories that are relevant to developmental and clinical research, and whether measurements continue to be accurate over long recording periods. Data processing algorithms are needed to classify the raw sensor data (i.e., linear and angular acceleration time series) into meaningful categories. Categorizing body position---supine on the back, prone on the belly, sitting, upright, or held off the ground by a caregiver---is a complex because movement can vary greatly \emph{within} a body position. An upright infant can be standing still or can be walking briskly across the room. A prone infant can be stationary in ``tummy time'', or they can crawl in a myriad of ways \autocite{AdolphVereijken1998}. Moreover, the configuration of the arms, legs, and torso within a body position can vary greatly in everyday contexts. Infants can sit on the floor in a tripod position with support from an arm, in a ``V'' position with legs fully extended, or in a ``W'' position with knees bent. Sitting on a caregiver's lap without the need to maintain balance means that the legs can dangle and the torso can lean in different directions. Sitting in a high chair or car seat reduces the magnitude of postural sway within sitting, and creates even more possibilities for how the arms and legs may move relative to the torso. Finally, caregivers frequently pick up and transport infants, creating motion signals that need to be differentiated from independent activity \autocite{KwonZavos2019,PatelShi2019}.

Modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults \autocite{Preece2009,Arif2015}, children \autocite{NamPark2013,RenDing2016,StewartNarayanan2018}, and infants \autocite{YaoPlotz2019,FranchakScott2021,AiraksinenRasanen2020}. Three prior investigations have used different machine learning techniques to categorize infant body position from IMUs towards the goal of collecting full-day data. \textcite{AiraksinenRasanen2020} tested 4- to 8-month-olds in a laboratory visit with a 4-sensor array (one on each thigh and one below each shoulder), and found 95\% accuracy in distinguishing between body position categories that crawling infants could perform on the floor (excluding times that infants were held by caregivers). Using a wider age range of 6-18 months, \textcite{FranchakScott2021} found 98\% accuracy (\emph{kappa} = 95\%) in a laboratory validation study with a 3-sensor array (ankle, knee, thigh on a single leg) in categorizing body position that included infants who could both crawl and walk and also included a category for caregiver holding. Most recently, \textcite{AiraksinenGallen2022} conducted a validation study of body position classification in either a home or clinic testing 4- to 19-month-olds with their 4-sensor system, refining their previous method to detect moments that infants were carried by caregivers. Classification accuracy did not vary between lab and home settings, and was generally high (95\%, \emph{kappa} = .93). Although all three studies yielded promising classification accuracy, accuracy was assessed in brief (15-60 minute) sessions supervised by a researcher, leaving the open question of how well body position classification will scale to testing across an entire day of natural home life.

@@

Accordingly, the overarching goal of the current study is to test the validity of long-form body position recording in the home during unsupervised, everyday behavior. Supervised recordings from past work \autocite{FranchakScott2021,AiraksinenRasanen2020,AiraksinenGallen2022}, whether in the home or in the lab, let researchers set up the situation to encourage or restrict certain behaviors. Prior work focused on ``free play'', in which caregivers were asked to play with the infant without restraining the infant or otherwise shaping how they could move. However, in a real day, non-play activities (e.g., eating lunch in a high chair) create challenging situations for applying automated classification of body position. For example different types of sitting---independently on the floor, supported on a caregiver's lap, or restrained in a high chair---all need to be scored as sitting. Moreover, classifiers must be able to detect new variants of behaviors that might arise over the course of a real day; it is impossible for researchers to gather training data for every possible variation that might occur. Thus, our central question is whether models trained on video-recorded observations at the beginning of the day generalize to predict behavior at a later time? Assessing the validity of temporally \emph{distal} periods---when infants and caregivers are unsupervised and free to follow their everyday routines---is a crucial step to establish whether automatic classification can be used to measure body position across a day.

@@

Infants were recruited in one of two age groups: \emph{Younger} infants were between 4 and 7 months and \emph{older} infants were between 11 and 14 months. There were 8 infants in the 4-7 month group and 14 in the 11-14 month group. Ten infants were female and 12 were male. Families were recruited through social media advertisements and from community events in Southern California. Most infants were reported to be either Hispanic and White or Non-Hispanic and White. Families were compensated \$30 for every home recording session they completed. The BLINDED Institutional Review Board reviewed and approved all procedures associated with the study. All caregivers gave their informed consent before the start of the study.

Most participants were tested in a single session (\emph{n} = 15), but 7 participants contributed between 2-4 sessions as part of an ongoing longitudinal study. Although including more data from some participants could over-represent their characteristics in the model, we reasoned that this drawback was outweighed by having more available data to use for training. Only 1 session was excluded due to a technical error---one of the four IMU sensors failed to record, resulting in an unusable set of data for classification. Across the two age groups, we report data on a total of 34 sessions, with 14 sessions from younger infants and 20 sessions from older infants. Across sessions, younger infants' age ranged from 3.8 to 7.2 months (\emph{M} = 5.2) and older infants' age ranged from 10.7 to 14.2 months (\emph{M} = 11.7). The total number of recording sessions (34) exceeded the number of sessions employed in comparable past work: 10 in \textcite{NamPark2013}, 15 in \textcite{FranchakScott2021}, 22 in \textcite{AiraksinenRasanen2020}, and 33 in \textcite{YaoPlotz2019}.

@@

Four inertial movement units (IMUs) were used to record infant movement across the day (MC10 Biostamp). A custom garment was designed to hold the IMUs: Internal pockets were sewn into snug-fitting infant leggings so that IMUs would stay close to the body (reducing vibration). A pocket over the thigh and a pocket just above the ankle were sewn on the lateral surface of the right and left legs of the garment. Multiple sizes of the garment were created (modeled on US 3-6 mo, 6-9 mo, 9-12 mo, and 12-18 mo sizing). Caregivers indicated in advance which size would fit their infant, and to ``size down'' if between sizes to ensure a snug fit and minimize sensor movement. Each garment had a distinct pattern on the seat of the pants so that caregivers could identify front versus back and place the garment in the correct orientation.

Each sensor had sufficient battery and onboard torage to record accelerometer and gyroscope data for approximately 12 hours. We chose a sampling rate of 62.5 Hz (one of the available presets) based on prior work that used rates of 50-64 Hz \autocite{FranchakScott2021,AiraksinenRasanen2020,YaoPlotz2019}. Infants also wore a LENA® recorder throughout the day in the front pocket of a LENA® shirt, located near the infant's chest, to determine whether data could be simultaneously recorded from the LENA® and IMU sensors (LENA® data were not analyzed in the current study).

@@

Because the camera only had the battery life to record for \textasciitilde3 hours (automatically split into two 90-minute video files), this divided the day into different periods for analysis. As seen in the bottom of Figure \ref{fig:exemplar-timeline}), the \emph{video period} comprised the first three hours of recording starting from the researcher's arrival when they turned on the camera. The first 90-minute video file, termed the \emph{proximal comparison}, contained the activities during the guided call followed by a period during which infants and caregivers resumed their normal activities. Because this video contained the synchronization point, the data in this period had high temporal synchrony between IMU and video data. Synchronization errors were estimated to be less than 30-60 ms (1-2 video frames). The second video file comprised the \emph{distal comparison}. This video recorded the next 90 minutes of natural activity. However, because of a limitation in the camera, there was a variable gap of \textasciitilde40 s between the two videos, so synchronization in the distal comparison video was coarser, with estimated temporal offsets of \textasciitilde5 s in either direction.

@@

The proximal and distal comparison videos for each participant were annotated by trained human coders to identify infant \emph{body position} into one of 5 mutually-exclusive categories following the definitions in prior work \autocite{FranchakScott2021}: supine, prone, sitting, upright, or held by caregiver. All coding was done using Datavyu software (datavyu.org). A Databrary repository contains the entire video recording, coding files, and raw IMU data for a single participant (\url{https://nyu.databrary.org/volume/1580}).

Supine was coded when the infant was lying on their back, on their side, or was reclined up to a 45 Prone was coded when the infant was lying on their stomach, was stationary supported by the hands/knees or the hands/feet, or was crawling. We scored sitting to include any form of the following seated positions: 1) infants sat with their buttocks on a surface, such as on the floor or a caregiver's lap, 2) infant was in a kneeling-sit position, in which their knees were on the ground with their legs tucked underneath the buttocks, and 3) infant was in a seating device, such as a high chair, that kept the torso oriented perpendicular to the ground (a reclined position, such as in a young infant's car seat, would be counted as supine). Upright was coded when the infant was standing or squatting on the ground with two feet or walking (regardless of whether infants' balance was assisted by a caregiver or with their hands holding onto something for support). Our goal in creating a category for ``held by caregiver'' was to separate times when infants were in control of their body position from times when they were suspended in the air (rather than resting on furniture or a surface). Held was coded when infants were carried off of the groun. However, when the caregiver was sitting with the infant in their lap the infant's body position was coded as if the caregiver was a surface (e.g., if the infant was sitting on the caregiver's lap this was coded as sitting). Times during the video when the infant was out of view were excluded. Periods when the sensor garment was adjusted or taken off the infant were also excluded, as were transitions between body positions.

@@

Classification training and prediction was conducted on a windowed dataset that summarized the raw, 62.5 Hz motion signals within 4-s windows. Data were reduced in time by creating overlapping moving windows (4-s long, comprising 250 samples) starting each second, which is a common unit of analysis in prior studies of human activity classification \autocite{FranchakScott2021,NamPark2013,AiraksinenRasanen2020}. For each 4-s window, we aggregated the 250 samples to create single scores for a variety of motion features---summary statistics that could be fed into the machine learning model. The minimum, maximum, 25th percentile, 75th percentile, mean, median, skew, kurtosis, standard deviation, and sum were computed for each signal (e.g., right thigh X linear acceleration, left ankle pitch angular acceleration). The 10 summary statistics and 24 sensor signals generated 240 columns of motion features that described movement within each window. Furthermore, a series of cross-sensor and cross-orientation summaries (such as the correlation, magnitude, and difference between pairs of sensors) added an additional 196 columns of motion features. The 436 total motion features corresponded to a single body annotation code for each 4-s window. Windows were only used for training/testing if they contained a single body position for \textgreater{} 75\% (3 s) of time within the window to ensure that motion signals could be linked to an unambiguous example of each behavior.

@@

\caption{Metrics of agreement between human-annotated body position and model predictions of body position from the proximal comparison period. Overall accuracy (A) and Cohen's Kappa (B) are plotted separately for group models and individual models. Each blue circle represents that accuracy for each recording session. Horizontal black bars indicate the mean across sessions.}\label{fig:metrics}
\end{figure}

@@

\begin{tabular}{lllllll}
\toprule
 & \multicolumn{3}{c}{Group} & \multicolumn{3}{c}{Individual} \\
\cmidrule(r){2-4} \cmidrule(r){5-7}
Sensors & Median & Mean & SD & Median & Mean & SD\\
\midrule
All & 0.894 & 0.846 & 0.131 & 0.933 & 0.916 & 0.072\\ \midrule
Left Hip/Ankle & 0.891 & 0.824 & 0.177 & 0.915 & 0.901 & 0.071\\
Right Thigh/Ankle & 0.842 & 0.800 & 0.167 & 0.897 & 0.898 & 0.086\\
Both Thighs & 0.859 & 0.791 & 0.195 & 0.907 & 0.898 & 0.072\\
Both Ankles & 0.797 & 0.719 & 0.196 & 0.864 & 0.844 & 0.108\\ \midrule
Left Thigh & 0.842 & 0.762 & 0.208 & 0.889 & 0.872 & 0.077\\
Left Ankle & 0.717 & 0.667 & 0.208 & 0.851 & 0.825 & 0.117\\
Right Thigh & 0.846 & 0.772 & 0.178 & 0.896 & 0.876 & 0.092\\
Right Ankle & 0.704 & 0.683 & 0.160 & 0.830 & 0.819 & 0.114\\
\bottomrule
\end{tabular}

@@

The first set of results showed that group and individual models trained from data during the proximal comparison period were within the proximal period. In the next analysis, we examine long-term performance by testing how accurately models trained from the proximal period could predict body position during the distal comparison period. A single group model was created using all sessions' proximal period training data (rather than group models leaving out a single session); the same individual models were used. Distal videos had only coarse temporal synchrony with motion recordings which precluded calculating accuracy based on the proportion of matching events. Instead, we summed the amount of time infants were predicted to be in each of the 5 body position categories from the model and compared that to the summed time for the body positions based on human coding \autocite{YaoPlotz2019,FranchakScott2021}.

@@

\caption{Age differences in daily body position predicted from (A) group models and (B) individual models. Each circle represents one full-day recording session's proportion of time in each body position (y-axis) against age in months (x-axis). Black lines indicate best fit regression lines indicating decreases in supine time and increases in sitting and upright time with age.}\label{fig:age}
\end{figure}
